---
title: "r-week-8-hw-eeb174"
author: "Dustin Wong"
date: "March 8, 2017"
output: html_document
---
```{r}
library(ggplot2)
setwd("/home/eeb177-student/Desktop/eeb-177/homework/week-8-hw-eeb174/")

canids <- read.csv("/home/eeb177-student/Desktop/eeb-177/homework/week-8-hw-eeb174/canid-ranges.csv", header =F, as.is = T)

names(canids) <- c("genus", "species", "minage", "maxage")
head(canids)
```

```{r}
canid_occ <- ggplot(canids, aes( species, ymin = maxage, ymax=minage, colour = genus))
canid_occ <- canid_occ + geom_linerange()
canid_occ <- canid_occ + theme(legend.position="none") #removes legend
canid_occ <- canid_occ + coord_flip() #flips axis/graph
canid_occ <- canid_occ +  theme(axis.text.y = element_text(size=4)) #larger text size
canid_occ <- canid_occ + theme(axis.ticks.y=element_blank()) #removes tick marks
canid_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) #removes white space on graphs
canid_occ <- canid_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))  #adds title/labels
canid_occ
```

What taxonomic group is displayed?
print("Canids")

How many unique fossil species are contained within it?
cat("Unique species:", nrow(canids))

How many unique fossil genera?
genera <- unique(canids$genus)
cat("Number of Genera:", length(genera))

What species has the longest time
canids$time <- canids$maxage - canids$minage
max_time_taxa<- which.max(canids$time) 
cat(canids[max_time_taxa,2],canids[max_time_taxa,5], "Ma"





## exploring data frames
#Challenge1
df <- data.frame(first = c('Grace'), last = c('Hopper'), lucky_number = c(0), stringsAsFactors = FALSE)
df <- rbind(df, list('Marie', 'Curie', 238))
df <- cbind(df, coffeetime = c(TRUE,TRUE))

#Challenge2
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv") gapminder <- read.csv(file = "data/gapminder-FiveYearData.csv")
source(file = "scripts/load-gapminder.R")

#Challenge3
The object gapminder is a data frame with columns, country and continent are factors, year is an integer vector, and pop, lifeExp, and gdpPercap are numeric vectors.



## subsetting data
#challenge1
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
1) x[c(2,3,4)]
2) x[2:4]
3) x[-c(1,5)]

#challenge 2
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
answer: x[-which(names(x) == "g")]
Only A and C are correct

#challenge 3
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
answer: x_subset <- x[x4]
print(x_subset)

#challenge 4
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
answer: D only

#challenge 5
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
answer: xlist$b[2]

#challenge 6
mod <- aov(pop ~ lifeExp, data=gapminder)
answer: attributes(mod) ## `df.residual` is one of the names of `mod`
mod$df.residual

#challenge 7
gapminder[gapminder$year = 1957,]
corrected: gapminder[gapminder$year == 1957,]

gapminder[,-1:4]
corrected: gapminder[,-c(1:4)]

gapminder[gapminder$lifeExp > 80]
corrected: gapminder[gapminder$lifeExp > 80,]

gapminder[1, 4, 5]
corrected: gapminder[1, c(4, 5)]

gapminder[gapminder$year == 2002 | 2007,]
corrected: gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
gapminder[gapminder$year %in% c(2002, 2007),]

#challenge 8
1) gapminder is a data.frame so needs to be subsetted on two dimensions. gapminder[1:20, ] subsets the data to give the first 20 rows and all columns.
2) gapminder_small <- gapminder[c(1:9, 19:23),]



## Creating Publication-Quality Graphics
#challenge 1
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()

#challenge2
ggplot(data = gapminder, aes(x = year, y = lifeExp, color=continent)) + geom_point()
#seems to be expected based on how rich/developed the region is

#challenge3
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) + geom_point() + geom_line(aes(color=continent))

#challenge4a&4b
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point(size=3, color="orange") + scale_x_log10() + geom_smooth(method="lm", size=1.5)

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + geom_point(size=3, pch=17) + scale_x_log10() + geom_smooth(method="lm", size=1.5)

#challenge5
ggplot(data = gapminder, aes(x = gdpPercap, fill=continent)) + geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()


